Trade-offs between sustainable development goals in systems of cities

### 2022-05-10 Response_to_reviewers.tex

    \bigskip

    \item \textit{The methods section is in many ways a black box with - alright 30 cities with a zipf law distribution, and a stepwise, 10-year interval progress horizon to tell the story of innovation spread, but why 30 cities, why rank-size distribution, etc., etc. the choices made are not discussed properly and the methodology is more or less of uncommented and left to references in works by others. I am NOT suggesting that the authors would benefit from going all formula in the methods section - but if there would have been a clear link between the introduction and the methods section, a motivation for the choices made would be stronger.}

    \medskip

    $\rightarrow$ The model description was further extended and mathematical equations added.

@@

    \bigskip

%Specific comments:

    \item \textit{The authors write: ``We consider the "innovation" SDG (goal 9) and the "climate" 192 SDG (goal 14) as conflicting objectives. … Empirical evidence does not suggest globally a simultaneous reduction of emissions through innovation (Chen 198 and Lee, 2020).'' Though I share the authors concern for environment and the role of innovation, there are several theories discussing the relationship between innovation and climate that are of opposing views. As the authors suggest - there may not be a support for a global simultaneous reduction of emissions (but why have the results both be global and simultaneous to be valid?), but examples where innovation is reducing emissions are well-known, and the scientific debate (though being highly criticized) about the environmental Kuznets curve is juxtaposing your view of the relationship between innovation and emissions. Examples of the later is that innovations have reduced emissions in vehicles (private transport in particular) but at the same time - the emission variable you are using as a constant (well as a function of distance and population growth) in your model - assumes that there is no innovation in transport. Under this assumption the results make sense but what is the purpose with that kind of a find?}

@@

\medskip

\begin{enumerate}
    \item \textit{On literature review. Seen from the paper's organization structure, literature review is included in the Introduction. Its benefit is to save space and to be concise. But for more logical organization, another literature review had better to be added.}

    \medskip

    $\rightarrow$ The paper was reorganised, in particular the introduction was expanded and split into subsections including different streams of literature review.

@@

    \bigskip

    \item \textit{On the model. In Section 2, for concise purpose, the model's mathematical description is omitted. But the mathematical description for the model is rather critical for understanding the idea of the paper as well as the specific model simulation followed. So, the mathematical description for the model should be given even in reduced way.}

    \medskip

    $\rightarrow$ Mathematical equations for model dynamics were added.

### 2022-05-10 paper.tex
Update on Overleaf.

\journal{Journal of Urban Management}

@@

\author{}
%\author[label1]{Juste Raimbault\corref{cor1}}
%\author[label2]{Denise Pumain}

%\address[label1]{CASA, University College London}
%\address[label2]{G{\'e}ographie-cit{\'e}s, Universit{\'e} Paris 1}

%\cortext[cor1]{Corresponding author: juste.raimbault@polytechnique.edu}

@@

The 17 Sustainable Development Goals that were defined by the United Nations in 2015 \citep{nations2015sustainable} are challenging the adaptation capability of many cities in the world. As they host a growing proportion of the world population and are responsible for a large part of greenhouse gas emissions (Christen 2014), cities are at the core of environmental transition policies \citep{romero2011cities}. Indeed, cities are explicitly mentioned in SDG 11 \citep{nations2015sustainable}: ``Make cities and human settlements inclusive, safe, resilient, and sustainable''. This citation demonstrates that many other SDGs involving a diversity of urban issues and stakeholders are already embedded in that specific goal \citep{vaidya2020sdg}.

In the ongoing adaptive process of the environmental transition, cities are not isolated. They have been engaged for long in a co-opetition process for attracting population and economic activities that have created many interdependencies in their evolution leading to interpret cities as “systems within systems of cities” (Berry, 1964). These systems of cities are an almost universal form of spatial organization of societies to inhabit the earth space. This organization has proven to be very sustainable for several millennia, because it offered a way to continuously adapt the political, economic and technological innovations that societies imagined over the centuries \citep{pumain2020theories}. That remarkable ability to resilience through self-renewal has expanded all over the world up to the point that the largest metropolises may have now become ``global rulers'' \citep{glaeser2020urban}. However, for the last few decades urban systems seem to be threatened by excessive emissions and overexploitation of the planet's energy and material resources \citep{nijkamp2014sustainable,kourtit2020global}. While demographic and economic growth have supported and accelerated the development of urbanization and the proliferation of systems of cities for two centuries, they are now accused of being the cause of climate change problems and the rise of social inequalities \citep{davis2006planet,glaeser2009inequality}. It seems that the excessive mobility associated with the fragmentation of the value chains of globalized production has progressively decoupled urban expansion from the proper management of planetary resources \citep{rozenblat2018urban}. It is unlikely that the technological solutions that are provided under the label of “smart cities” could easily solve these problems \citep{caragliu2011smart,kourtit2020global}, especially because of the huge diversity of cities internal layouts and their already established networks operating internal and external interactions \citep{}(Caruso et al. 2022).

Regarding the adaptation process to SDGs, recent research follows two main lines, involving theoretical and empirical investigations.  On the theoretical side, the last three decades have produced a huge literature, mainly oriented toward assessing the relationship between gas emissions and GDP over time. The environmental Kuznets curve (EKC) hypothesis proposes that there is an inverted U-shape relation between environmental degradation and income per capita \citep{dinda2004environmental, stern2004rise}. This has been taken to imply that economic growth will eventually redress the environmental impacts observed from the early stages of economic development. However, difficulties linked with conceptual definitions and indicators measurements, although helping to clarify the issue, have not led to a complete consensus about the robustness of that hypothesis \citep{harbaugh2002reexamining}. Moreover, most tests regarding that hypothesis have been made at a macrogeographic level and compare national statistical evolutions. The observations made at city level are more recent and fragmented. Empirical investigations about SDGs in cities still lack of well-established and comparable sources of information. Thus they follow a diversity of directions: either identifying relevant indicators for benchmarking and monitoring the improvement process \citep{giles2020achieving}, or measuring their variations according to city size \citep{laituri2021sdg} and according to specific practices and specific national contexts (Mudau et al., 2020 for inclusivity South Africa, Mohan et al. 2021 for road safety in India, Koch and Krellenberg 2018 for Germany, Zhu et al 2022 for the relationship between carbon emissions and polycentric structures in China).

@@

@@

We propose in this paper to study trade-offs between different SDGs in systems of cities. We consider systems of cities at the macroscopic scale, and more particularly the dynamics of innovation diffusion and population growth. Using a stylised model for such urban dynamics, we apply a bi-objective optimisation genetic algorithm, to explore how trade-offs can occur in such systems.

@@

We work with a stylised model for the dynamics of urban systems at the macroscopic scale (i.e. a country or a continent or any integrated region of the world). This model is based on innovation diffusion dynamics and their impact on population growth. It was first formulated by \cite{favaro2011gibrat}, within the context of an evolutionary urban theory \citep{pumain1997pour}. A similar agent-based model was used to explore assumptions on the emergence of systems of cities themselves \citep{schmitt2015half}. A modified version was described by \cite{raimbault2020model} as an urban evolution model, including an urban genome shared and mutated across cities. As this particular version can furthermore be setup on stylised systems of cities, we use it in our multi-objective optimisation approach.

We describe the model below, but without detailed equations for the purpose of staying concise. For a full mathematical description of the model, the reader is referred to \cite{raimbault2020model}.

@@

The simulated urban system is composed by cities, which location in the geographical space and number are fixed in time. They are characterised at each time step by their population $P_i (t)$ and by adoption rates by their populations for different innovations (urban genome $\delta_{i,c} (t)$).

@@

Synthetic systems of cities are generated with random locations, an initial rank-size hierarchy which can be tuned (otherwise fixed to a default value of 1, to mirror a Zipf law distribution for city size \citep{cottineau2017metazipf}), and a number of 30 cities. The largest city has initially a population of 100,000 and the model is run for 50 time steps.

@@

Starting from the initial state, the model updates population and innovation step by step. At each time step (of an order of magnitude of 10 years - the effects are observed on long time scales), the following procedure is used:

\begin{enumerate}
	\item innovations are diffused between cities using a spatial interaction model - innovations with a higher utility will diffuse more quickly and obtain higher adoption shares \citep{hagerstrand1968innovation};
	\item populations are updated following an other spatial interaction model \citep{raimbault2020indirect}, with a population growth advantage for cities being more innovative;
	\item new innovations may be invented in cities, following a probability determined by a mutation rate and by population (with a given hierarchy across cities, following the empirical fact of superlinear innovation scaling \citep{arbesman2009superlinear});
	\item if a new innovation emerges, it has an initial penetration share fixed by one parameter, and a utility randomly distributed (normal or log-normal law), with a fixed standard deviation and an average corresponding to the current empirical average of existing innovation utilities.
\end{enumerate}
% note: do not necessarily have an increase of utility: in avg should stay the same with normal, deviate to higher with log-normal?
%  -> higher utilities are selected with diffusion, so in practice should always increase in expectancy - this may be an internal validation aspect to be checked

@@

\begin{enumerate}
    \item spatial interaction range for innovation diffusion
    \item spatial interaction range for population growth
    \item mutation (innovation) rate
    \item level of hierarchy to select cities inventing a new innovation (scaling of innovations with regard to urban size)
    \item rate of early adopters
    \item standard-deviation of new innovation utilities
    \item type of distribution for new innovation utilities.
\end{enumerate}

@@

Model optimisation is achieved by integrating the model into the OpenMOLE platform \cite{reuillon2013openmole}. This free and open source software facilitates model embedding into a workflow system, distribution of computation into high-performance computing infrastructures, and provides a simple access to state-of-the-art model validation methods.

@@

To conclude, we have provided a first stylised insight into trade-offs between SDGs in systems of cities at the macroscopic scale, which can be applied from a theoretical viewpoint to validate or unvalidate urban theories, and be used as a basis towards more practical application towards sustainable long-term territorial policies \cite{rozenblat2018conclusion}.

% additional dimensions? wage inequalities \cite{shutters2022urbanization}

>>>>

\journal{Journal of Urban Management}
%\journal{}

@@

%\author{}
\author[label1,label2,label3]{Juste Raimbault\corref{cor1}}
\author[label4]{Denise Pumain}

\address[label1]{LASTIG, Univ Gustave Eiffel, IGN-ENSG}
\address[label2]{CASA, University College London}
\address[label3]{UPS CNRS 3611 ISC-PIF}
\address[label4]{UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Universit{\'e} Paris 1}

\cortext[cor1]{Corresponding author: juste.raimbault@polytechnique.edu}

@@

\subsection{Urban systems and urban sustainability}

The 17 Sustainable Development Goals that were defined by the United Nations in 2015 \citep{nations2015sustainable} are challenging the adaptation capability of many cities in the world. As they host a growing proportion of the world population and are responsible for a large part of greenhouse gas emissions \citep{christen2014atmospheric}, cities are at the core of environmental transition policies \citep{romero2011cities}. Indeed, cities are explicitly mentioned in SDG 11 \citep{nations2015sustainable}: ``Make cities and human settlements inclusive, safe, resilient, and sustainable''. This citation demonstrates that many other SDGs involving a diversity of urban issues and stakeholders are already embedded in that specific goal \citep{vaidya2020sdg}.

In the ongoing adaptive process of the environmental transition, cities are not isolated. They have been engaged for long in a co-opetition process for attracting population and economic activities that have created many interdependencies in their evolution leading to interpret cities as ``systems within systems of cities'' \citep{berry1964cities}. These systems of cities are an almost universal form of spatial organization of societies to inhabit the earth space. This organization has proven to be very sustainable for several millennia, because it offered a way to continuously adapt the political, economic and technological innovations that societies imagined over the centuries \citep{pumain2020theories}. That remarkable ability to resilience through self-renewal has expanded all over the world up to the point that the largest metropolises may have now become ``global rulers'' \citep{glaeser2020urban}. However, for the last few decades urban systems seem to be threatened by excessive emissions and overexploitation of the planet's energy and material resources \citep{nijkamp2014sustainable,kourtit2020global}. While demographic and economic growth have supported and accelerated the development of urbanization and the proliferation of systems of cities for two centuries, they are now accused of being the cause of climate change problems and the rise of social inequalities \citep{davis2006planet,glaeser2009inequality}. It seems that the excessive mobility associated with the fragmentation of the value chains of globalized production has progressively decoupled urban expansion from the proper management of planetary resources \citep{rozenblat2018urban}. It is unlikely that the technological solutions that are provided under the label of ``smart cities'' could easily solve these problems \citep{caragliu2011smart,kourtit2020global}, especially because of the huge diversity of cities internal layouts and their already established networks operating internal and external interactions \citep{caruso2022no}.

\subsection{Investigating transitions towards sustainability}

Regarding the adaptation process to SDGs, recent research follows two main lines, involving theoretical and empirical investigations.  On the theoretical side, the last three decades have produced a huge literature, mainly oriented toward assessing the relationship between gas emissions and GDP over time. The environmental Kuznets curve (EKC) hypothesis proposes that there is an inverted U-shape relation between environmental degradation and income per capita \citep{dinda2004environmental, stern2004rise}. This has been taken to imply that economic growth will eventually redress the environmental impacts observed from the early stages of economic development. However, difficulties linked with conceptual definitions and indicators measurements, although helping to clarify the issue, have not led to a complete consensus about the robustness of that hypothesis \citep{harbaugh2002reexamining}. Moreover, most tests regarding that hypothesis have been made at a macrogeographic level and compare national statistical evolutions. The observations made at city level are more recent and fragmented. Empirical investigations about SDGs in cities still lack of well-established and comparable sources of information. Thus they follow a diversity of directions: either identifying relevant indicators for benchmarking and monitoring the improvement process \citep{giles2020achieving}, or measuring their variations according to city size \citep{laituri2021sdg} and according to specific practices and specific national contexts, such as for example inclusivity in South Africa \citep{mudau2020assessment}, road safety in India \citep{mohan2021future}, contextualisation of goals in Germany \citep{koch2018contextualize}, or the relationship between carbon emissions and polycentric structures in China \citep{zhu2022did}.

@@

\subsection{Conflicting sustainability objectives}

@@

\subsection{Proposed approach}

We propose in this paper to study trade-offs between different SDGs in systems of cities. Our research question couples the two streams of literature reviewed above: we inquire how urban dynamics stylised modeling can be applied to the exploration of conflicting SDGs dimensions. We consider systems of cities at the macroscopic scale, and more particularly the dynamics of innovation diffusion and population growth. Using a stylised model for such urban dynamics, we apply a bi-objective optimisation genetic algorithm, to explore how trade-offs can occur in such systems.

@@

We work with a stylised model for the dynamics of urban systems at the macroscopic scale (i.e. a country or a continent or any integrated region of the world). This model is based on innovation diffusion dynamics and their impact on population growth. It was first formulated by \cite{favaro2011gibrat}, within the context of an evolutionary urban theory \citep{pumain1997pour}. A similar agent-based model was used to explore assumptions on the emergence of systems of cities themselves \citep{schmitt2015half}. A modified version was described by \cite{raimbault2020model} as an urban evolution model, including an urban genome shared and mutated across cities. As this particular version can furthermore be setup on stylised systems of cities, we use it in our multi-objective optimisation approach. We give below a detailed description of model setup and dynamics.

@@

The simulated urban system is composed by cities, which location in the geographical space and number are fixed in time. They are characterised at each time step by their population $P_i (t)$ and by adoption rates by their populations for different innovations, what corresponds to an ``urban genome'' as a matrix $\delta_{i,c} (t)$ with $c$ being the index for successive innovations.

@@

Synthetic systems of cities are generated with random locations, an initial rank-size hierarchy which can be tuned (otherwise fixed to a default value of 1, to mirror a Zipf law distribution for city size \citep{cottineau2017metazipf}), and a number of 30 cities. The largest city has initially a population of 100,000 and the model is run for 50 time steps. These values correspond to an order of magnitude of a regional or national system of cities on a period of one or two centuries, which is the correct application context for this model \citep{favaro2011gibrat}. Such a setup is furthermore usual for synthetic cases in this family of models, such as for the SimpopNet model systematic exploration \citep{raimbault2020unveiling}, for a co-evolution model between cities and transportation networks \citep{raimbault2021modeling}, and for this specific version of the innovation diffusion model \citep{raimbault2020model}.

@@

Starting from the initial state, the model updates population and innovation step by step. At each time step (of an order of magnitude of a few years years - the effects are observed on long time scales), the following procedure is used:

\begin{enumerate}
	\item innovations are diffused between cities using a spatial interaction model - innovations with a higher utility will diffuse more quickly and obtain higher adoption shares \citep{hagerstrand1968innovation}; innovation shares are updated, with $p_{c,i,t} = \delta_{c,i,t} \cdot \frac{P_{i}(t)}{\sum_k P_k (t)}$ the city-level share of adoption, $u_c$ the utility of innovation $c$, and $d_I$ the spatial interaction range for innovation diffusion, following
        \begin{equation}
            \delta_{c,i,t} = \frac{\sum_j p_{c,j,t-1}^{\frac{1}{u_c}} \cdot \exp{(-\frac{d_{ij}}{d_I})}}{\sum_c \sum_j p_{c,j,t-1}^{\frac{1}{u_c}} \cdot \exp{(-\frac{d_{ij}}{d_I})}}
        \end{equation}
	\item populations are updated following an other spatial interaction model \citep{raimbault2020indirect}, with a population growth advantage for cities being more innovative; more precisely, new populations are computed as 
    	\begin{equation}
	        P_i(t) - P_i(t-1) = w_I\cdot \sum_j \frac{V_{ij}}{<V_{ij}>}
	    \end{equation}
	where the spatial interaction potential for population growth is given by 
	    \begin{equation}
            V_{ij}= \frac{P_{i}(t-1) \cdot P_{j}(t-1)}{(\sum_k P_k(t-1))^2} \cdot \exp{\left(-\frac{d_{ij}}{d_G} \cdot \prod_c \delta_{c,i,t}^{\phi_{c,t}}\right)}
        \end{equation}
    with $\phi_{c,t} = \sum_i \delta_{i,c,t}\cdot P_i(t-1) /\sum_{i,c} \delta_{i,c,t}\cdot P_{i}(t-1)$ macroscopic adoption rate, $w_I$ population growth rate parameter , and $d_G$ spatial interaction range for population growth;
	\item new innovations may be invented in cities, following a probability determined by a mutation rate and by population with a given hierarchy across cities as
	    \begin{equation}
	        p = \beta \cdot \left(P_i (t) / \max_k P_k (t)\right)^{\alpha_I}
	    \end{equation}
	with $\beta$ intrinsic innovation rate and $\alpha_I$ innovation hierarchy; this follows the empirical fact of superlinear innovation scaling \citep{arbesman2009superlinear};
	\item if a new innovation emerges, it has an initial penetration share fixed by one parameter $r_0$, and a utility randomly distributed (normal or log-normal law, with the type of law being also a model parameter), with a fixed standard deviation $\sigma$ and an average corresponding to the current empirical average of existing innovation utilities.
\end{enumerate}
% note: do not necessarily have an increase of utility: in avg should stay the same with normal, deviate to higher with log-normal?
%  -> higher utilities are selected with diffusion, so in practice should always increase in expectancy - this may be an internal validation aspect to be checked

@@

\begin{enumerate}
    \item spatial interaction range for innovation diffusion $d_I$
    \item spatial interaction range for population growth $d_G$
    \item mutation (innovation) rate $\beta$
    \item level of hierarchy to select cities inventing a new innovation (scaling of innovations with regard to urban size) $\alpha_I$
    \item rate of early adopters $r_0$
    \item standard-deviation of new innovation utilities $\sigma$
    \item type of distribution for new innovation utilities.
\end{enumerate}

@@

Model optimisation is achieved by integrating the model into the OpenMOLE platform \citep{reuillon2013openmole}. This free and open source software facilitates model embedding into a workflow system, distribution of computation into high-performance computing infrastructures, and provides a simple access to state-of-the-art model validation methods.

@@

Adding supplementary optimisation dimensions, to include further SDGs in simulated dynamics, is also an important future model development.
% additional dimensions? wage inequalities \cite{shutters2022urbanization}

Finally, several model limitations would require investigation.

To conclude, we have provided a first stylised insight into trade-offs between SDGs in systems of cities at the macroscopic scale, which can be applied from a theoretical viewpoint to validate or invalidate urban theories, and be used as a basis towards more practical application towards sustainable long-term territorial policies \citep{rozenblat2018conclusion}.
